The promise of community-driven preprints in ecology and evolution

### 2024-05-09 ms/ms.qmd
Some changes to intro

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. While preprint use in biology has had a slower uptake [@Berg2016], in ecology and evolution, new discipline-specific preprint servers, such as *EcoEvoRxiv*, provide a means by which ecologists and evolutionary biologists can disseminate research findings. Preprints are expected to promote open science, increase the visibility of research (e.g., [@fu2019releasing]), and facilitate the rapid sharing of new discoveries that can have significant impacts on fundamental and applied knowledge globally. Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings which can lead to more citations [e.g., @colavizza2024analysis]; 2) helping establish precedence of research findings; 3) removing financial barriers to publication; and 4) encouraging constructive and critical feedback from the research community [@Proulx2013; @bourne2017ten; @vale2015accelerating].

Preprint servers can empower researchers to make their findings more accessible, open, and transparent but only if they are used as forums for spreading and discussing findings within a research community. Significant barriers to preprint use still exist globally – from a lack of clarity around preprint policies at journals [@Klebel2020] to a stigma that preprints are viewed by the research community to be of poor quality [@Chiarelli2019] (but see [@Faser2022]). Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution. Such an understanding can help inform future initiatives and allow us to work harder in further breaking down barriers to scientific publishing.

>>>>

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. While preprint use in biology has had a slower uptake [@Berg2016], in ecology and evolution, new discipline-specific preprint servers, such as *EcoEvoRxiv*, provide a means by which ecologists and evolutionary biologists can disseminate research findings. Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings which can lead to more citations [e.g., @colavizza2024analysis]; 2) helping establish precedence of research findings; 3) removing financial barriers to publication; and 4) encouraging constructive and critical feedback from the research community [@Proulx2013; @bourne2017ten; @vale2015accelerating]. Preprints are expected to promote open science, increase the visibility of research (e.g., [@fu2019releasing]), and facilitate the rapid sharing of new discoveries that can have significant impacts on fundamental and applied knowledge globally.

Preprint servers can empower researchers to make their research findings more accessible, open, and transparent but only if they are used as forums for spreading and discussing findings within a research community. Significant barriers to preprint use still exist globally – from a lack of clarity around preprint policies at journals [@Klebel2020], to a stigma that preprints are viewed by the research community to be of poor quality [@Chiarelli2019] (but see [@Faser2022]). Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution. Such an understanding can help inform future initiatives and allow us to work harder in further breaking down barriers to scientific publishing.
